---
title: "covid 19"
author: "Logan DeHay"
date: "10/11/2021"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
options(digits=2)
require(tidyverse)
require(tigerstats)
require(lubridate)
kRdata <- read.csv(file = 'kRdata.csv')
```
### Data is from the Patient_Info dataset from South Korea 1 Jan to 1 May 
### This is contained in a downloadable zip file from kaggle

Some of the data was uncoded for sex - sex must be just male or female
77 of the data entries were coded sex  as blank, so we omit these.
state of a patient must be died or alive - we adjust that as well   
  
# Question 1 - Is dying independent of sex




Table of patients:
```{r}

# cross tabulation, patient state by gender

with(kRdata, table(Sex, result))

```

## We create two bar charts - one based on frequency and the other on percent, use ggplot commands to do this

`
```{r}
barchartGC(~result + Sex,data=kRdata)
barchartGC(~result + Sex,data=kRdata, type="percent")

```
  
Both populations have a minor sample who contracted the disease, and according to the bar charts above, the males generally die more than the females. However, there are more females in the data than males. Despite this data size difference, the data suggests an increased probability of dying from COVID in Korea if you are male.
  



  
## Numerical results
```{r}
table1 <- xtabs(~result + Sex, data=kRdata)
rowPerc(table1)
colPerc(table1)
```
  
The table above displays that 56% of the alive population are females while 36% of the dead population are female. The second table further illustrates that 3% of the total sample population of males are dead from COVID while only 1.4% of the females are dead from COVID in Korea.
  
## Inferential results 
```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
```
  
Looking at the test above, the counts when compared to the counts expected by the null hypothesis are vastly different. Our data displayed that 44 males died as opposed to 25 females even though there is a greater sample of females. This indicates a correlation in the data between death and sex, disproving the null hypothesis. As was shown in the Fisher exact test, the odds ratio of death for male to females is 2.3:1. This means that if you are a male you are 2.3 times more likely to die than a female. This is very different form the expected null ratio of 1:1. Taking into account the 95% confidence interval, the test suggests that the ratio can range from 1.4:1 to 3.9:1. Finally, the calculated p-value is very small, which illustrates that we can reject the null hypothesis and assert that sex has a significant effect on death from COVID in Korea.
  
## Conclusions
  
We began this process by determining the type of helper file necessary in relation to the data we had. Our data is composed of largely categorical variables, indicating the need for a Categorical~Categorical helper file to compare these variables together. To compare, we utilized bar charts and tables as visual displays of the data and as a result, could  conclude that more males died from COVID-19 in Korea than females despite the larger sample size of females. To further conclude if there is a correlation between sex and death by COVID, we utilized a Fisher exact test for count data and a chi squared test. Both p-values were sufficiently small enough to suggest that the probability of the data results being unbiased is extremely unlikely, and as a result, that sex has a significant impact on death by COVID.
  