---
title: "covid 19"
author: "Logan DeHay"
date: "10/11/2021"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
options(digits=2)
require(tidyverse)
require(tigerstats)
require(lubridate)
kRdata <- read.csv(file = 'kRdata.csv')
```
### Data is from the Patient_Info dataset from South Korea 1 Jan to 1 May 
### This is contained in a downloadable zip file from kaggle

Some of the data was uncoded for sex - sex must be just male or female
77 of the data entries were coded sex  as blank, so we omit these.
state of a patient must be died or alive - we adjust that as well   
  
# Question 1 - Is dying independent of sex




Table of patients:
```{r}

# cross tabulation, patient state by gender

with(kRdata, table(Sex, result))

```

## We create two bar charts - one based on frequency and the other on percent, use ggplot commands to do this

`
```{r}
barchartGC(~result + Sex,data=kRdata)
barchartGC(~result + Sex,data=kRdata, type="percent")

```
  
Both populations have a minor sample who contracted the disease, and according to the bar charts above, the males generally die more than the females. However, there are more females in the data than males. Despite this data size difference, the data suggests an increased probability of dying from COVID in Korea if you are male.
  



  
## Numerical results
```{r}
table1 <- xtabs(~result + Sex, data=kRdata)
rowPerc(table1)
colPerc(table1)
```
  
The table above displays that 56% of the alive population are females while 36% of the dead population are female. The second table further illustrates that 3% of the total sample population of males are dead from COVID while only 1.4% of the females are dead from COVID in Korea.
  
## Inferential results 




## Conclusions

